---
title: "PS5"
author: "Justin Post"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Read in some data

Let's start by reading in some data from a URL. This is a `.csv` file so it is a comma delimited file. Opening it, we saw that there is a header row and some missing values. But it seems like it should read in just fine with `read_csv()`
```{r}
library(tidyverse)
sleep_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/sleep_data.csv")
```

The file at <https://www4.stat.ncsu.edu/~online/datasets/sleep_data_info.txt> gives the reference for where this data comes from. Let's recreate the first plot (almost) that we see in that paper.

```{r}
#create a plot and define parts of the plot
ggplot(sleep_data, aes(x = TotalSleepTime, 
                       y = term_gpa,
                       color = cohort)) +
  geom_point(color = "blue", alpha = 0.1, aes(size = midpoint_sleep)) + #add a scatter plot layer
  geom_smooth(method = "lm", se = FALSE) + #add SLR fits
  xlab("Total Sleep Time") + #change x label
  ylab("Spring Term GPA") + #change y label
  ggtitle("Scatter Plot with SLR Fits") #change title
```



```{r}
#create a plot and define parts of the plot
ggplot(sleep_data, aes(x = TotalSleepTime, 
                       y = term_gpa,
                       color = cohort)) +
  geom_point(alpha = 0.4, aes(size = midpoint_sleep, shape = cohort)) + #add a scatter plot layer
  geom_smooth(method = "lm", se = FALSE) + #add SLR fits
  xlab("Total Sleep Time") + #change x label
  ylab("Spring Term GPA") + #change y label
  ggtitle("Scatter Plot with SLR Fits") + #change title
  scale_size(range = c(0.5, 3), guide = "legend")
```

## Numeric Summaries


```{r}
colnames(sleep_data)
sleep_data %>%
  group_by(cohort) %>%
  summarize(meanTotSleep = mean(TotalSleepTime),
            sdTotSleep = sd(TotalSleepTime))

# same kind of summaries for all dbl variables in the data set
sleep_data <- sleep_data %>%
  mutate(demo_race_cat = factor(demo_race),
         demo_gender_cat = factor(demo_gender)) 

sleep_data %>%
  group_by(cohort) %>%
  summarize(across(where(is.double) & !c(demo_race, demo_gender, subject_id), 
                list(mean, sd)))

sleep_data %>%
  group_by(cohort) %>%
  summarize(across(where(is.double) & !c(demo_race, demo_gender, subject_id), 
                list(mean = mean, sd = sd), 
                .names = "{.col}.{.fn}"))
```




