---
title: "PS6"
author: "Justin Post"
date: "2023-09-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises from *Advanced R*

Exercise 3 from section 11.1.2.

Use both a `for` loop and `lapply()` to fit linear modesl to the `mtcars` data frame using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1/disp),
  mpg ~ disp + wt,
  mpg ~ I(1/disp) + wt
)
```

First, let's learn a bit about fitting a basic linear model (multiple linear regression) in R. This can be done with the `lm()` function. 

```{r}
fit <- lm(mpg ~ disp, data = mtcars)
```

We can obtain a summary of the model fit using `summary()`.

```{r}
summary(fit)
```

We can obtain diagnostic plots (for checking assumptions about the model) using the generic `plot()` function. This actually calls `plot.lm()` (see `methods(plot)`, which leads to `getAnywhere("plot.lm")` to get the source code for this method!).
```{r}
class(fit)
str(fit, max.level = 1)
plot(fit)
```

If we wanted to get a scatter plot with the fitted line overlayed. We can do that using `Base R` or the `tidyverse`.

```{r}
plot(x = mtcars$disp, y = mtcars$mpg)
#sequence for plotting the model fit
x <- seq(from = 50, to = 500, by = 0.5)
#predictions come from the generic predict() function (we'll cover all this in topic 3!)
preds <- predict(fit, newdata = data.frame(disp = x))
lines(x = x, y = preds)

library(tidyverse)
ggplot(mtcars, aes(x = disp, y = mpg)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

Now let's use a `for` loop to fit the model for each formula. Here we could use `seq_along(formulas)` to get an indexing vector. I tend to use the `1:length(formulas)` kind of syntax. (You could also directly iterate over the formulas elements!)

```{r}
#use a for loop to fit our models
seq_along(formulas)
for (i in 1:length(formulas)){
  lm(formulas[[i]], data = mtcars) %>%
    summary() %>%
    print()
}

#to save the output of the lm() fits, we can initialize a list
fit_list <- list()
for (i in 1:length(formulas)){
  fit_list[[i]] <- lm(formulas[[i]], data = mtcars) 
}
```

This is a bit easier using `lapply()`. Note that we pass the name of the function only, `lm`, and not a **call** to the function `lm(x, ...)`. As we have an additional argument to `lm()` other than the formulas, we can pass that as an additional argument to `lapply()`.

```{r}
#use lapply instead
fit_objects <- lapply(X = formulas, FUN = lm, data = mtcars)
fit_objects[[1]]
str(fit_objects, max.level = 1)
```

Now let's fit our models but also pull out some useful stats.

```{r}
#useful stats to get out
fit
sum_fit <- summary(fit)
class(summary(fit))
str(sum_fit)
#pull out coefficients
sum_fit$coefficients
str(sum_fit$coefficients)
dimnames(sum_fit$coefficients)[[1]]
#pull out adjusted r squared
sum_fit$adj.r.squared
```

Great, let's write a custom function to do the fit and to pull out those values. We'll return them all in a list.

```{r}
fit_objects <- lapply(X = formulas, FUN = function(x){
  fit <- lm(x, data = mtcars)
  sum_fit <- summary(fit)
  coefs <- sum_fit$coefficients
  adj_r_sq <- sum_fit$adj.r.squared
  list(fit = fit, coefs = coefs, adj_r_sq = adj_r_sq)
})

str(fit_objects[[1]], max.level = 1)
```

Alternatively, we can supply a `my_data` argument to our custom function that allows us to pass in a different data frame (as done earlier).

```{r}
fit_objects <- lapply(X = formulas, 
                      FUN = function(x, my_data){
                        fit <- lm(x, data = my_data)
                        sum_fit <- summary(fit)
                        coefs <- sum_fit$coefficients
                        adj_r_sq <- sum_fit$adj.r.squared
                        list(fit = fit, coefs = coefs, adj_r_sq = adj_r_sq)
                      }, 
                      my_data = mtcars)
```

We could also save our function as an object and call it in a similar manner to how we called `lm()`.

```{r}
custom_fun <- function(x, my_data){
                        fit <- lm(x, data = my_data)
                        sum_fit <- summary(fit)
                        coefs <- sum_fit$coefficients
                        adj_r_sq <- sum_fit$adj.r.squared
                        list(fit = fit, coefs = coefs, adj_r_sq = adj_r_sq)
                      }

fit_objects <- lapply(X = formulas, 
                      FUN = custom_fun, 
                      my_data = mtcars)
```


To give an example where we might iterate over a data frame instead, we'll apply the `lm()` function to each column of the `mtcars` data frame using `mpg` as the response.

```{r}
#create a matrix of predictors (similar to a data frame)
predictors <- mtcars[, -1]
#use apply to apply a custom function to each column of the data frame
slr_fits <- apply(X = predictors, MARGIN = 2, FUN = function(x1){
  lm(mtcars$mpg ~ x1)
})
```







